A ``Dream'' Interface Design

### 2014-11-26 dream-design/dream-design.tex
final touches

However, the literature suggests that overall feedback is still insufficient and remains a major usability issue with touchscreen devices. \cite{Tinwala:2010:ETE:18	68914.1868972, Kane:2011:UGB:1978942.1979001, Hardy:2008:TIT:1409240.1409267, El-Glaly:2013:TTF:2460625.2460665, Buxton:1986:HID:22339.22386}. A study by Hussain Tinwala and Scott MacKenzie submits that the lack of physical keys requires heightened visual attention from the user, which diverts the user's concentration from the thoughts being expressed \cite{Tinwala:2010:ETE:18 68914.1868972}. The lack of physical keys not only diverts the attention of some users, but it also makes the device  almost entirely unusable for other users. A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI) since these users are only able to develop a spatial mental model for the touchscreen through ``dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.

@@

My ``dream'' interface design incorporates functionality of a product that is currently being developed by Tactus Technology \cite{Tactus}. Tactus Technology, a company based in Fremont, California, creates technology that can reshape the surface of touchscreens on demand. The technology allows real physical buttons to dynamically appear and disappear into a flat touchscreen (see Figure~\ref{tactus1}). Small fluid channels are routed throughout the Tactile Layer and enable fluid to expand the top polymer layer to create the physical buttons \cite{Tactus}.

Although this technology from Tactus is extremely bleeding-edge, the prototypes have already received a fair amount of recognition \cite{CNN, I-Zone, PCMag, Wired}. Reviewers have articulated how the technology seems to be ``downright magical'' \cite{CNN}. Though the look and feel of Tatctus technology seems totally futuristic, the technology is already beginning to appear in consumer devices. In 2013, Touch Revolution, the largest-volume glass projected capacitive multi-touch screen manufacturer in the world, announced a partnership with Tactus Technology \cite{TactusAvailability}. The technologies from the two companies are being combined and manufactured into consumer devices today \cite{TactusAvailability}. Not only is this technology becoming more and more available, but it is becoming more customizable as well. Companies will soon be able to customize the panel for different types of buttons, say for example, the buttons on a TV remote \cite{CNN}.

@@

My ``dream'' user interface design makes use of the varied tactile feedback that this technology from Tactus can provide.

@@

\subsection{Optional Physical Grid}
The design incorporates Tactus technology in order to provide an optional tactile grid for the device (see Figure~\ref{wireframe-grid}). The grid layout would consist of a set of very thin, three-dimensional horizontal and vertical lines. If the grid is enabled by the user, the set of horizontal and vertical lines appear on the screen. If the grid is disabled by the user, the lines recede. The horizontal and vertical lines are arranged with equal space between each line. The main purpose of this grid is to help IBSVI formulate a mental reference for location awareness. The fact that the grid can easily be disabled makes this feature unobtrusive for users who do not find benefit from it.

\begin{figure}[ht]
\centering
\includegraphics[width=2in]{wireframe-grid.png} 
\caption{A simple wireframe demonstrating the concept of the grid layout. Each line in the figure would actually be a physical line that appears on the device's screen with the use of Tactus's microfluidic technology.}
\label{wireframe-grid}
\end{figure}

\subsection{``Intelligent'' Tactile Buttons}
The interface design also includes tactile buttons that use Tactus's microfluidic technology to rise up and recede in an ``intelligent'' manner (see Figure~\ref{buttons}). Ideally, only buttons applicable to the input of the user or current state of the device would appear at any given time. For example, in an instance where the user is provided with a binary yes/no choice, only two tactile buttons should appear---one for yes and one for no. Similarly, if all possible buttons were applicable for a given situation, then all possible buttons should appear.

@@

\subsection{Aural Feedback}
\label{aural-feedback}
The design of this interface was created in such a way that the tactile aspects are intended to operate in conjunction with other technologies that provide aural feedback. Since the grid and tactile buttons can only help the user know \textit{where} an application is, this aural feedback can inform the user of exactly \textit{what} an application is.

@@

\section{Usage Scenarios}
There exist several usage scenarios for this interface design. One such scenario is when an Individual with Blindness of Servere Visual Impairment (IBSVI) attempts to locate an application on a touchscreen mobile device. Sighted users employ vision to figure out what and where to touch. Since IBSVI cannot do this, typical touchscreen devices lack accessibility for IBSVI. As previously touched on in Section~\ref{aural-feedback}, a typical solution to this problem is to use voice over technology to provide information about what is on the screen at any given time. However, there is nothing to guide IBSVI to the location of a specific application in the first place. Since there are no landmarks to support navigation for the IBSVI, the user must use dead-reckoning and trial-and-error methods in order to find applications specified by the voice over technology.

@@

\section{Rationale}
The driving force behind the creation and development of this interface design is feedback. Feedback is considered a critical concept in the field of interaction design. Donald Norman, the author of \textit{The Design of Everyday Things}, repeatedly articulates the importance of showing the effect of an action \cite{Norman02}. Without feedback, the user is always wondering whether anything has happened. In my ``dream'' interface design, the tactile feedback from the physical buttons serve to help the user understand when an action has been performed. Furthermore, the additional aural feedback of the design also assists in communicating the current state of the device to the user.

Another interaction design concept that Norman discusses is how systems with more options and less feedback can be very difficult to use \cite{Norman02}. My design includes ``intelligent'' physical buttons so that the device provides the user with the minimum number of options any given time. Since the buttons will only appear in situations when they are needed, it narrows the options of the user. This ultimately decreases the cognitive load of the user as they have less buttons from which to choose.

This layout is designed so that IBSVI can engage their spatial cognition, perception and sensing resources while interacting with touch screens. The design will only enable the grid-like system on-command so to not distract or annoy those who do not benefit from it.

Feedback. In design, it is important to show the effect of an action. Without feedback, one is always wondering whether anything has happened. Maybe the button wasn't pushed hard enough; maybe the machine has stopped working; maybe it is doing the wrong thing. Without feedback, we turn equipment off at improper times or restart unnecessarily, losing all our recent work. Or we repeat the command and end up having the operation done twice, often to our detriment. Feedback is critical.

Usability research evaluating touchscreen efficacy for both IBSVI and sighted users
still needs to be covered in the literature. I completely agree that the lack of accessibility
for IBSVI is a major issue with touchscreen devices. However, providing more tactile
feedback can increase usability not just for IBSVI, but also for sighted users. In Donald
Norman’s book, The Design of Everyday Things, Norman touches on this principle of
feedback. Norman describes how telephones in the past “were designed with much
more care and concern for the user” as “the push buttons were designed to give an
appropriate feel—tactile feedback.” [20] Norman asserts that modern telephone systems
are “so difficult to use” since they have “more features and less feedback.” [20] As tactile
feedback is significant for IBSVI and sighted users alike, research should examine how
both groups of users respond to proposed solutions. I think that this could possibly
lead to a solution that becomes more widely accepted.

\section{Usability Metric ``Forecast''}
Predict to see an increase in learnability for sighted users and IBSVI alike. Should be more efficient for IBSVI but may be slightly less efficient for sighted users. They can no longer press a button quite as easily. Now that buttons are physical, they will have to lift their finger slightly more and press down slightly harder. This should not take any major hits with efficiency. Rate of errors should be less for both parties.
\clearpage

>>>>

Although these alternative aural and haptic signals are provided for the user, the literature suggests that overall feedback is still insufficient and remains a major usability issue with touchscreen devices. \cite{Tinwala:2010:ETE:18	68914.1868972, Kane:2011:UGB:1978942.1979001, Hardy:2008:TIT:1409240.1409267, El-Glaly:2013:TTF:2460625.2460665, Buxton:1986:HID:22339.22386}. A study by Hussain Tinwala and Scott MacKenzie submits that the lack of tactile feedback requires heightened visual attention from the user, which diverts the user's concentration from the thoughts being expressed \cite{Tinwala:2010:ETE:18 68914.1868972}. The lack of tactile feedback not only diverts the attention of some users, but it also makes the device  almost entirely unusable for other users. A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI). These users are only able to develop a spatial mental model for the touchscreen through ``dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.

@@

This ``dream'' interface design incorporates functionality of a product that is currently being developed by Tactus Technology \cite{Tactus}. Tactus Technology, a company based in Fremont, California, creates technology that can reshape the surface of touchscreens on demand. The technology allows real physical buttons to dynamically appear and disappear into a flat touchscreen (see Figure~\ref{tactus1}). Small fluid channels are routed throughout the Tactile Layer and enable fluid to expand the top polymer layer to create the physical buttons \cite{Tactus}.

Although this technology from Tactus is extremely bleeding-edge, the prototypes have already received a fair amount of recognition \cite{CNN, I-Zone, PCMag, Wired}. Reviewers have articulated how the technology seems to be ``downright magical'' \cite{CNN}. Though the look and feel of Tatctus technology seems totally futuristic, the technology is already beginning to appear in consumer devices. In 2013, Touch Revolution, the largest-volume glass projected capacitive multi-touch screen manufacturer in the world, announced a partnership with Tactus Technology \cite{TactusAvailability}. The technologies from the two companies are being combined and manufactured into consumer devices today \cite{TactusAvailability}. Not only is this technology becoming more  available, but it is becoming more customizable as well. Companies will soon be able to customize the panel for different types of buttons, say for example, the buttons on a TV remote \cite{CNN}.

@@

This ``dream'' user interface design makes use of the varied tactile feedback that the technology from Tactus can provide.

@@

\subsection{Optional Physical Grid}
The design incorporates Tactus technology in order to provide an optional tactile grid for the device (see Figure~\ref{wireframe-grid}). The grid layout would consist of a set of very thin, three-dimensional horizontal and vertical lines. If the grid is enabled by the user, the set of horizontal and vertical lines appears on the screen. If the grid is disabled by the user, the lines recede. The horizontal and vertical lines are arranged with equal space between each line. The main purpose of this grid is to help IBSVI formulate a mental reference for location awareness. The fact that the grid can easily be disabled makes this feature unobtrusive for users who do not find benefit from it.

\begin{figure}[ht]
\centering
\includegraphics[width=2in]{wireframe-grid.png} 
\caption{A simple wireframe demonstrating the concept of the grid layout. Each line in the figure would actually be a physical line that appears on the device's screen with the use of Tactus' microfluidic technology.}
\label{wireframe-grid}
\end{figure}

\subsection{``Intelligent'' Tactile Buttons}
The interface design also includes tactile buttons that utilize Tactus' microfluidic technology in order to rise up and recede in an ``intelligent'' manner (see Figure~\ref{buttons}). Ideally, only buttons applicable to the input of the user or current state of the device would appear at any given time. For example, in an instance where the user is provided with a binary yes/no choice, only two tactile buttons should appear---one for yes and one for no. Similarly, if all possible buttons were applicable for a given situation, then all possible buttons should appear.

@@

\subsection{Aural Feedback}
\label{aural-feedback}
The design of this interface was created in such a way that the tactile aspects can operate in conjunction with other technologies that provide aural feedback. Since the grid and tactile buttons can only help the user know \textit{where} an application is, this aural feedback can inform the user of exactly \textit{what} an application is.

@@

\section{Usage Scenarios}
There exist several usage scenarios for this interface design. One such scenario is when an Individual with Blindness of Severe Visual Impairment (IBSVI) attempts to locate an application on a touchscreen mobile device. Sighted users employ vision to figure out what and where to touch. Since IBSVI cannot do this, typical touchscreen devices lack accessibility for IBSVI. As previously touched on in Section~\ref{aural-feedback}, a typical solution to this problem is to use voice over technology to provide information about what is on the screen at any given time. However, there is nothing to guide IBSVI to the location of a specific application in the first place. Since there are no landmarks to support navigation for the IBSVI, the user must use dead-reckoning and trial-and-error methods in order to find applications specified by the voice over technology.

@@

\section{Rationale}
The driving force behind the creation and development of this interface design is feedback. Feedback is considered a critical concept in the field of interaction design. Donald Norman, the author of \textit{The Design of Everyday Things}, repeatedly articulates the importance of showing the effect of an action \cite{Norman02}. Without feedback, the user is always wondering whether or not anything has happened. In this ``dream'' interface design, the tactile feedback from the physical buttons serves to help the user understand when an action has been performed. Furthermore, the additional aural feedback of the design also assists in communicating the current state of the device to the user.

Another interaction design concept that Norman discusses is how systems with more options and less feedback can be very difficult to use \cite{Norman02}. My design includes ``intelligent'' physical buttons so that the device provides the user with the minimum number of options any given time. Since the buttons will only appear in situations when they are needed, it narrows the options for the user. This ultimately decreases the cognitive load on the user as they have less buttons from which to choose.

This design was also created in a way that allows IBSVI to engage their spatial cognition, perception and sensing resources while interacting with touchscreen mobile devices. A previous study from Virginia Polytechnic Institute and State University demonstrates how touchscreen devices with some kind of a physical overlay typically improve learnability and efficiency while decreasing error rates for IBSVI \cite{El-Glaly:2013:TTF:2460625.2460665}. The optional grid of this ``dream'' interface is very similar to the physical overlay from the study, and should thus improve usability metrics for IBSVI with touchscreen phones.

\section{Usability Metric ``Forecast''}
If implemented then tested, I would expect this interface to have better usability metrics with respect to learnability than conventional touchscreen mobile devices.  More physical landmarks allow all users to engage their spatial cognition more fully when interacting with touchscreens. Furthermore, evidence in the past has demonstrated  that more tactile feedback can make a system more learnable \cite{El-Glaly:2013:TTF:2460625.2460665, conf/chi/KaneWL11}.

Studies in the past have also suggested that more tactile feedback makes IBSVI more efficient when using the system \cite{El-Glaly:2013:TTF:2460625.2460665, Buxton:1986:HID:22339.22386}. This interface might, however, be slightly less efficient for users who do not have visual impairments. These users will have to lift their fingers more and press down significantly harder with the tactile buttons. I would expect this difference in efficiency to be negligible though.

I would also expect all users to be more perform less errors as well as be more satisfied with this ``dream'' interface. As tactile buttons are harder to press, there would be less likelihood that one would be pressed accidentally. Furthermore, studies have shown that most users enjoy tactile feedback \cite{Freeman:2014:TFA:2663204.2663280}.

\clearpage
